Tropical Andes Functional diversity for plants and Frugivores

### 2024-06-24 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
create functions to calculate and map FDis, make multi-panel plots for maps

# Load required packages
```{r}
library(mFD)
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
library(rlang)
library(doParallel)
library(foreach)
```

# Set file paths
```{r}
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

@@

# Create function
```{r}

@@

```{r}
# Extract cell values
frugivore_cellFDis_100km <- spatial_fdis_frugivore_grid
hist(frugivore_cellFDis_100km$fdis_value)
```

@@

```{r}
plantgridFDisTA_100km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    ggtitle("100 km") +
    scale_fill_distiller(palette = "Greens", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
plantgridFDisTA_100km
```

```{r}
# Extract cell values
plant_cellFDis_100km <- spatial_fdis_plant_grid
hist(plant_cellFDis_100km$fdis_value)
```

## 75 km

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_plant <- rowSums(PAM_plant_site_final_75km)
subset_matrix_plant <- PAM_plant_site_final_75km[row_sums_plant >= 4, ]
```

```{r}
# Generate coordinates
subset_coords_plant <-site_loc_key_plant_75km[rowSums(PAM_plant_site_final_75km) >= 4, ]
subset_coords_plant_sp <-subset_coords_plant[,1:2 ]

plant_fd_sp_75km <- data.frame(subset_coords_plant_sp, fdis_plant)
```

```{r}
# Convert the dataframe to sf format
spatial_fdis_plant_75km <- st_as_sf(plant_fd_sp_75km, coords = c("Longitude", "Latitude"))
```

```{r}
# set crs of sf objects
spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>% st_set_crs(5389)
```

```{r}
# Extract coordinates 
spatial_fdis_plant_coords_75km <- st_coordinates(spatial_fdis_plant_75km)

# Add coordinates as separate columns
spatial_fdis_plant_75km <- spatial_fdis_plant_75km %>%
  mutate(x = spatial_fdis_plant_coords_75km[, 1], y = spatial_fdis_plant_coords_75km[, 2])
```

```{r}
summary(spatial_fdis_plant_75km)
```

```{r}
TAGrid__75km <- TApoly %>%
  st_make_grid(cellsize = c(75000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())

spatial_fdis_plant_75km$cellid <- st_intersects(spatial_fdis_plant_75km, TAGrid_75km)$nn

spatial_fdis_plant_grid_75km <- TAGrid_75km %>%
   st_join(spatial_fdis_plant_75km) %>%
   group_by(cellid) %>%
   summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
   ungroup()
```

```{r}
summary(spatial_fdis_plant_grid_75km)
```

```{r}
plantgridFDisTA_75km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_plant_grid_75km, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    ggtitle("100 km") +
    scale_fill_distiller(palette = "Greens", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
plantgridFDisTA_75km
```

>>>>

# Load required packages
```{r}
library(mFD)
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
library(rlang)
library(doParallel)
library(foreach)
library(purrr)
library(ggpubr)
```

# Set file paths
```{r}
data_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

@@

@@

```{r}
ggsave("frugivoregridFDisTA_100km.png", frugivoregridFDisTA_100km, path = figure_path)
# Extract cell values
frugivore_cellFDis_100km <- spatial_fdis_frugivore_grid
hist(frugivore_cellFDis_100km$fdis_value)
```

## Create function to calculate FD as fdis for frugivores for other spatial grains
Uses inputs from trait objects created above to avoid redundancies 
```{r}
calculate_fd_as_fdis_frugivores <- function(PAM_site_final) {
  # Calculate row sums and subset matrix early to avoid unnecessary calculations
  row_sums_frugivore <- rowSums(PAM_site_final)
  subset_matrix_frugivore <- PAM_site_final[row_sums_frugivore >= 4, ]

  # Select the necessary columns from sp_faxes_coord_frugivore
  sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[, c("PC1", "PC2", "PC3", "PC4")]

  ## check number of species names
  print(nrow(sp_faxes_coord_frugivore_sub))
  print(ncol(subset_matrix_frugivore))

  sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
  row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names

  #remove NAs
  sp_faxes_coord_frugivore_sub <- na.omit(sp_faxes_coord_frugivore_sub)
  subset_matrix_frugivore <- na.omit(subset_matrix_frugivore)

  print(nrow(sp_faxes_coord_frugivore_sub))
  print(ncol(subset_matrix_frugivore))

  if (nrow(sp_faxes_coord_frugivore_sub) != ncol(subset_matrix_frugivore)) {
    stop("Number of rows in sp_faxes_coord_frugivore_sub does not match number of columns in subset_matrix_frugivore")
  }

  sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
  subset_matrix_frugivore <- as.matrix(subset_matrix_frugivore)

  # Use three fewer cores than available
  num_cores <- parallel::detectCores() - 3

  # Set up parallel backend
  cl <- makeCluster(num_cores)
  registerDoParallel(cl)

  tryCatch({
    # Define chunk size
    chunk_size <- ceiling(nrow(subset_matrix_frugivore) / num_cores)

    # Split data into chunks for parallel processing
    chunks <- split(seq_len(nrow(subset_matrix_frugivore)), ceiling(seq_len(nrow(subset_matrix_frugivore)) / chunk_size))

    # Process chunks in parallel
    results_list <- foreach(chunk_indices = chunks, .combine = rbind, .packages = "mFD") %dopar% {
      subset_matrix_frugivore_chunk <- subset_matrix_frugivore[chunk_indices, , drop = FALSE]

      alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
        sp_faxes_coord = sp_faxes_coord_frugivore_sub, 
        asb_sp_w = subset_matrix_frugivore_chunk, 
        ind_vect = "fdis", 
        details_returned = TRUE
      )
     # Extract fdis values and match with corresponding cell numbers
      fdis_values <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
      cell_numbers <- rownames(subset_matrix_frugivore)[chunk_indices]

      # Create a data frame for this chunk
      chunk_results <- data.frame(Cell_Number = cell_numbers, fdis = fdis_values)
      chunk_results
    }

    # Combine all chunk results into a single data frame
    all_results <- do.call(rbind, results_list)

    # Stop the parallel backend
    stopCluster(cl)

    return(results_list)
  }, error = function(e) {
    stop("Error in parallel processing: ", conditionMessage(e))
  })
}

```{r}
fdis_frugivore_75km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_75km)
```

```{r}
fdis_frugivore_50km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_50km)
```

```{r}
fdis_frugivore_25km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_25km)
```

```{r}
fdis_frugivore_10km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_10km)
```

```{r}
fdis_frugivore_5km <- calculate_fd_as_fdis_frugivores(PAM_frugivore_site_final_5km)
```

## Map FD for spatial grains other than 100 km
```{r}
mapping_fd_as_fdis_frugivores <- function(PAM_frugivore_site, site_loc_key_frugivore, fdis_frugivore, resolution_meters) {
  # Generate coordinates
  subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site) >= 4, ]
  subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

  frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)

  # Convert the dataframe to sf format
  spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude", "Latitude"))

  # set crs of sf objects
  spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(5389)

  # Extract coordinates 
  spatial_fdis_frugivore_coords <- st_coordinates(spatial_fdis_frugivore)

  # Add coordinates as separate columns
  spatial_fdis_frugivore <- spatial_fdis_frugivore %>%
    mutate(x = spatial_fdis_frugivore_coords[, 1], y = spatial_fdis_frugivore_coords[, 2])

  TAGrid <- TApoly %>%
    st_make_grid(cellsize = c(resolution_meters)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())

  spatial_fdis_frugivore$cellid <- st_intersects(spatial_fdis_frugivore, TAGrid)$nn

  spatial_fdis_frugivore_grid <- TAGrid %>%
   st_join(spatial_fdis_frugivore) %>%
   group_by(cellid) %>%
   summarize(fdis_value = sum(fdis_frugivore, na.rm = TRUE)) %>%
   ungroup()

  frugivoregridFDisTA <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
  frugivoregridFDisTA

  list(frugivoregridFDisTA = frugivoregridFDisTA,
         spatial_fdis_frugivore_grid = spatial_fdis_frugivore_grid)
}
```

```{r}
mapping_75km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_75km, site_loc_key_frugivore_75km, fdis_frugivore_75km$fdis, 75000)
frugivoregridFDisTA_75km <- mapping_75km$frugivoregridFDisTA
frugivore_cellFDis_75km <- mapping_75km$spatial_fdis_frugivore_grid
```

```{r}
# add title to maps
(frugivoregridFDisTA_75km <- frugivoregridFDisTA_75km +
    ggtitle("75 km")
)
ggsave("frugivoregridFDisTA_75km.png", frugivoregridFDisTA_75km, path = figure_path)
```

```{r}
mapping_50km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_50km, site_loc_key_frugivore_50km, fdis_frugivore_50km$fdis, 50000)
frugivoregridFDisTA_50km <- mapping_50km$frugivoregridFDisTA
frugivore_cellFDis_50km <- mapping_50km$spatial_fdis_frugivore_grid
```

```{r}
# add title to maps
(frugivoregridFDisTA_50km <- frugivoregridFDisTA_50km +
    ggtitle("50 km")
)
ggsave("frugivoregridFDisTA_50km.png", frugivoregridFDisTA_50km, path = figure_path)
```

```{r}
mapping_25km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_25km, site_loc_key_frugivore_25km, fdis_frugivore_25km$fdis, 25000)
frugivoregridFDisTA_25km <- mapping_25km$frugivoregridFDisTA
frugivore_cellFDis_25km <- mapping_25km$spatial_fdis_frugivore_grid
```

```{r}
# add title to maps
(frugivoregridFDisTA_25km <- frugivoregridFDisTA_25km +
    ggtitle("25 km")
)
ggsave("frugivoregridFDisTA_25km.png", frugivoregridFDisTA_25km, path = figure_path)
```

```{r}
mapping_10km <- mapping_fd_as_fdis_frugivores(PAM_frugivore_site_final_10km, site_loc_key_frugivore_10km, fdis_frugivore_10km$fdis, 10000)
frugivoregridFDisTA_10km <- mapping_10km$frugivoregridFDisTA
frugivore_cellFDis_10km <- mapping_10km$spatial_fdis_frugivore_grid
```

```{r}
# add title to maps
(frugivoregridFDisTA_10km <- frugivoregridFDisTA_10km +
    ggtitle("10 km")
)
ggsave("frugivoregridFDisTA_10km.png", frugivoregridFDisTA_10km, path = figure_path)
```

```{r}
# Generate coordinates
subset_coords_frugivore_5km <-site_loc_key_frugivore_5km[rowSums(PAM_frugivore_site_final_5km) >= 4, ]
subset_coords_frugivore_sp_5km <-subset_coords_frugivore_5km[,1:2 ]

frugivore_fd_sp_5km <- data.frame(subset_coords_frugivore_sp_5km, fdis_frugivore_5km)
```

```{r}
# Convert the dataframe to sf format
spatial_fdis_frugivore_5km <- st_as_sf(frugivore_fd_sp_5km, coords = c("Longitude", "Latitude"))
```

```{r}
# set crs of sf objects
spatial_fdis_frugivore_5km <- spatial_fdis_frugivore_5km %>% st_set_crs(5389)
```

```{r}
# Extract coordinates 
spatial_fdis_frugivore_coords_5km <- st_coordinates(spatial_fdis_frugivore_5km)

# Add coordinates as separate columns
spatial_fdis_frugivore_5km <- spatial_fdis_frugivore_5km %>%
  mutate(x = spatial_fdis_frugivore_coords_5km[, 1], y = spatial_fdis_frugivore_coords_5km[, 2])
```

```{r}
TAGrid_5km <- TApoly %>%
  st_make_grid(cellsize = c(5000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())

spatial_fdis_frugivore_5km$cellid <- st_intersects(spatial_fdis_frugivore_5km, TAGrid_5km)$nn

```{r}
# Determine number of cores to use
num_cores <- detectCores() - 2

# Register parallel backend
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Ensure the cluster stops when done
on.exit(stopCluster(cl))

# Split TAGrid_5km indices into chunks for parallel processing
chunk_indices <- split(seq_len(nrow(TAGrid_5km)), rep(1:num_cores, each = ceiling(nrow(TAGrid_5km) / num_cores), length.out = nrow(TAGrid_5km)))

# Perform spatial join and aggregation in parallel
results_list <- foreach(indices = chunk_indices, .combine = 'c', .packages = c("sf", "dplyr")) %dopar% {
  grid_chunk <- TAGrid_5km[indices, ]

  # Perform spatial join
  joined <- st_join(grid_chunk, spatial_fdis_frugivore_5km)

  # Group by cellid and summarize fdis values
  summarized <- joined %>%
    group_by(cellid) %>%
    summarize(fdis_value = sum(fdis, na.rm = TRUE)) %>%
    ungroup()

  # Ensure that summarized is a data frame and add debugging information
  summarized_df <- as.data.frame(summarized)

  # Debugging: Check the structure of the summarized data frame
  message("Processed chunk: ", indices[1], " to ", indices[length(indices)])
  str(summarized_df)

  list(summarized_df)  # Return as list to avoid unintended simplification
}

# Flatten the results_list and ensure each element is a data frame
flattened_results <- map_dfr(results_list, ~ as.data.frame(.x))

# Check the structure of the combined result
str(flattened_results)

# Convert to sf object if applicable
if ("geometry" %in% colnames(flattened_results)) {
  spatial_fdis_frugivore_grid <- st_as_sf(flattened_results)
} else {
  spatial_fdis_frugivore_grid <- flattened_results
}

# Check the structure of the final output
str(spatial_fdis_frugivore_grid)

```{r}
frugivoregridFDisTA_5km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
  frugivoregridFDisTA_5km
frugivore_cellFDis_5km <- spatial_fdis_frugivore_grid
```

```{r}
# add title to maps
(frugivoregridFDisTA_5km <- frugivoregridFDisTA_5km +
    ggtitle("5 km")
)
ggsave("frugivoregridFDisTA_5km.png", frugivoregridFDisTA_5km, path = figure_path)
```

@@

```{r}
plantgridFDisTA_100km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    ggtitle("100 km") +
    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,7.5)) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
plantgridFDisTA_100km
```

```{r}
ggsave("plantgridFDisTA_100km.png", plantgridFDisTA_100km, path = figure_path)
# Extract cell values
plant_cellFDis_100km <- spatial_fdis_plant_grid
hist(plant_cellFDis_100km$fdis_value)
```

## Create function to calculate FD as fdis for plants for other spatial grains
Uses inputs from trait objects created above to avoid redundancies 
```{r}
calculate_fd_as_fdis_plants <- function(PAM_site_final) {
  # Calculate row sums and subset matrix early to avoid unnecessary calculations
  row_sums_plant <- rowSums(PAM_site_final)
  subset_matrix_plant <- PAM_site_final[row_sums_plant >= 4, ]

  # Select the necessary columns from sp_faxes_coord_plant
  sp_faxes_coord_plant_sub <- sp_faxes_coord_plant[, c("PC1", "PC2", "PC3", "PC4")]

  if (nrow(sp_faxes_coord_plant_sub) != ncol(subset_matrix_plant)) {
    stop("Number of rows in sp_faxes_coord_plant_sub does not match number of columns in subset_matrix_plant")
  }

  sp_faxes_coord_plant_sub <- as.matrix(sp_faxes_coord_plant_sub)
  subset_matrix_plant <- as.matrix(subset_matrix_plant)

  # Use three fewer cores than available
  num_cores <- parallel::detectCores() - 3

  # Set up parallel backend
  cl <- makeCluster(num_cores)
  registerDoParallel(cl)

  tryCatch({
    # Define chunk size
    chunk_size <- ceiling(nrow(subset_matrix_plant) / num_cores)

    # Split data into chunks for parallel processing
    chunks <- split(seq_len(nrow(subset_matrix_plant)), ceiling(seq_len(nrow(subset_matrix_plant)) / chunk_size))

    # Process chunks in parallel
    results_list <- foreach(chunk_indices = chunks, .combine = rbind, .packages = "mFD") %dopar% {
      subset_matrix_plant_chunk <- subset_matrix_plant[chunk_indices, , drop = FALSE]

      alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
        sp_faxes_coord = sp_faxes_coord_plant_sub, 
        asb_sp_w = subset_matrix_plant_chunk, 
        ind_vect = "fdis", 
        details_returned = TRUE
      )
     # Extract fdis values and match with corresponding cell numbers
      fdis_values <- alpha_fd_indices_plant$functional_diversity_indices$fdis
      cell_numbers <- rownames(subset_matrix_plant)[chunk_indices]

      # Create a data frame for this chunk
      chunk_results <- data.frame(Cell_Number = cell_numbers, fdis = fdis_values)
      chunk_results
    }

    # Combine all chunk results into a single data frame
    all_results <- do.call(rbind, results_list)

    # Stop the parallel backend
    stopCluster(cl)

    return(results_list)
  }, error = function(e) {
    stop("Error in parallel processing: ", conditionMessage(e))
  })
}

```{r}
fdis_plant_75km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_75km)
```

```{r}
fdis_plant_50km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_50km)
```

```{r}
fdis_plant_25km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_25km)
```

```{r}
fdis_plant_10km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_10km)
```

```{r}
fdis_plant_5km <- calculate_fd_as_fdis_plants(PAM_site_final = PAM_plant_site_final_5km)
```

## Map FD for spatial grains other than 100 km
```{r}
mapping_fd_as_fdis_plants <- function(PAM_plant_site, site_loc_key_plant, fdis_plant, resolution_meters) {
  # Generate coordinates
  subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site) >= 4, ]
  subset_coords_plant_sp <- subset_coords_plant[, 1:2]

  plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)

  # Convert the dataframe to sf format
  spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude", "Latitude"))

  # set crs of sf objects
  spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(5389)

  # Extract coordinates 
  spatial_fdis_plant_coords <- st_coordinates(spatial_fdis_plant)

  # Add coordinates as separate columns
  spatial_fdis_plant <- spatial_fdis_plant %>%
    mutate(x = spatial_fdis_plant_coords[, 1], y = spatial_fdis_plant_coords[, 2])

  # Use three fewer cores than available
  num_cores <- detectCores() - 3
  cl <- makeCluster(num_cores)
  registerDoParallel(cl)

  on.exit(stopCluster(cl)) # Ensure the cluster is stopped when the function exits

  tryCatch({
    # Generate the grid
    TAGrid <- TApoly %>%
      st_make_grid(cellsize = resolution_meters) %>%
      st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
      st_cast("MULTIPOLYGON") %>%
      st_sf() %>%
      mutate(cellid = row_number())

    # Check if TAGrid is empty
    if (nrow(TAGrid) == 0) {
      stop("Error: No valid grid cells after intersection.")
    }

    # Divide the indices of the grid into chunks for parallel processing
    chunk_indices <- split(seq_len(nrow(TAGrid)), rep(1:num_cores, each = ceiling(nrow(TAGrid) / num_cores), length.out = nrow(TAGrid)))

    # Process chunks in parallel
    results_list <- foreach(indices = chunk_indices, .combine = rbind, .packages = c("sf", "dplyr")) %dopar% {
      chunk <- TAGrid[indices, ]

      # Perform the intersection
      intersected <- st_intersection(spatial_fdis_plant, chunk)

      # Calculate cellid for intersected
      intersected$cellid <- st_intersects(intersected, chunk)$nn

      # Ensure that intersected is a data frame
      as.data.frame(intersected)
    }

    # Combine all chunk results into a single data frame
    spatial_fdis_plant <- do.call(rbind, results_list)

    # Ensure the combined result is a data frame
    spatial_fdis_plant <- as.data.frame(spatial_fdis_plant)

    # Debugging: Check if 'cellid' column exists
    if (!"cellid" %in% colnames(spatial_fdis_plant)) {
      stop("Error: 'cellid' column not found in spatial_fdis_plant.")
    }

    # Summarize fdis values per grid cell
    spatial_fdis_plant_grid <- spatial_fdis_plant %>%
      group_by(cellid) %>%
      summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
      right_join(as.data.frame(TAGrid), by = "cellid")

    # Convert the grid to sf
    spatial_fdis_plant_grid <- st_as_sf(spatial_fdis_plant_grid)

    # Create the plot
    plantgridFDisTA <- ggplot() +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
      geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
      labs(fill = "FDis") +
      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0, 7.5)) +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      theme(panel.background = element_rect(fill = "lightblue"),
            text = element_text(size = 12),
            plot.title = element_text(hjust = 0.5),
            axis.text.x = element_text(angle = 45, hjust = 1))

    list(plantgridFDisTA = plantgridFDisTA, spatial_fdis_plant_grid = spatial_fdis_plant_grid)
  }, error = function(e) {
    message("Error in parallel processing: ", conditionMessage(e))
    NULL
  })
}
```

```{r}
mapping_75km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_75km, site_loc_key_plant_75km, fdis_plant_75km$fdis, 75000)
plantgridFDisTA_75km <- mapping_75km$plantgridFDisTA
plant_cellFDis_75km <- mapping_75km$spatial_fdis_plant_grid
```

```{r}
# add title to maps
(plantgridFDisTA_75km <- plantgridFDisTA_75km +
    ggtitle("75 km")
)
ggsave("plantgridFDisTA_75km.png", plantgridFDisTA_75km, path = figure_path)
```

```{r}
mapping_50km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_50km, site_loc_key_plant_50km, fdis_plant_50km$fdis, 50000)
plantgridFDisTA_50km <- mapping_50km$plantgridFDisTA
plant_cellFDis_50km <- mapping_50km$spatial_fdis_plant_grid
```

```{r}
# add title to maps
(plantgridFDisTA_50km <- plantgridFDisTA_50km +
    ggtitle("50 km")
)
ggsave("plantgridFDisTA_50km.png", plantgridFDisTA_50km, path = figure_path)
```

```{r}
mapping_25km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_25km, site_loc_key_plant_25km, fdis_plant_25km$fdis, 25000)
plantgridFDisTA_25km <- mapping_25km$plantgridFDisTA
plant_cellFDis_25km <- mapping_25km$spatial_fdis_plant_grid
```

```{r}
# add title to maps
(plantgridFDisTA_25km <- plantgridFDisTA_25km +
    ggtitle("25 km")
)
ggsave("plantgridFDisTA_25km.png", plantgridFDisTA_25km, path = figure_path)
```

```{r}
mapping_10km <- mapping_fd_as_fdis_plants(PAM_plant_site_final_10km, site_loc_key_plant_10km, fdis_plant_10km$fdis, 10000)
plantgridFDisTA_10km <- mapping_10km$plantgridFDisTA
plant_cellFDis_10km <- mapping_10km$spatial_fdis_plant_grid
```

```{r}
# add title to maps
(plantgridFDisTA_10km <- plantgridFDisTA_10km +
    ggtitle("10 km")
)
ggsave("plantgridFDisTA_10km.png", plantgridFDisTA_10km, path = figure_path)
```

```{r}
# Generate coordinates
subset_coords_plant_5km <-site_loc_key_plant_5km[rowSums(PAM_plant_site_final_5km) >= 4, ]
subset_coords_plant_sp_5km <-subset_coords_plant_5km[,1:2 ]

plant_fd_sp_5km <- data.frame(subset_coords_plant_sp_5km, fdis_plant_5km)
```

```{r}
# Convert the dataframe to sf format
spatial_fdis_plant_5km <- st_as_sf(plant_fd_sp_5km, coords = c("Longitude", "Latitude"))
```

```{r}
# set crs of sf objects
spatial_fdis_plant_5km <- spatial_fdis_plant_5km %>% st_set_crs(5389)
```

```{r}
# Extract coordinates 
spatial_fdis_plant_coords_5km <- st_coordinates(spatial_fdis_plant_5km)

# Add coordinates as separate columns
spatial_fdis_plant_5km <- spatial_fdis_plant_5km %>%
  mutate(x = spatial_fdis_plant_coords_5km[, 1], y = spatial_fdis_plant_coords_5km[, 2])
```

```{r}
summary(spatial_fdis_plant_5km)
```

```{r}
TAGrid_5km <- TApoly %>%
  st_make_grid(cellsize = c(5000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())

spatial_fdis_plant_5km$cellid <- st_intersects(spatial_fdis_plant_5km, TAGrid_5km)$nn

```{r}
# Determine number of cores to use
num_cores <- detectCores() - 2

# Register parallel backend
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Ensure the cluster stops when done
on.exit(stopCluster(cl))

# Split TAGrid_5km indices into chunks for parallel processing
chunk_indices <- split(seq_len(nrow(TAGrid_5km)), rep(1:num_cores, each = ceiling(nrow(TAGrid_5km) / num_cores), length.out = nrow(TAGrid_5km)))

# Perform spatial join and aggregation in parallel
results_list <- foreach(indices = chunk_indices, .combine = 'c', .packages = c("sf", "dplyr")) %dopar% {
  grid_chunk <- TAGrid_5km[indices, ]

  # Perform spatial join
  joined <- st_join(grid_chunk, spatial_fdis_plant)

  # Group by cellid and summarize fdis values
  summarized <- joined %>%
    group_by(cellid) %>%
    summarize(fdis_value = sum(fdis_plant, na.rm = TRUE)) %>%
    ungroup()

  # Ensure that summarized is a data frame and add debugging information
  summarized_df <- as.data.frame(summarized)

  # Debugging: Check the structure of the summarized data frame
  message("Processed chunk: ", indices[1], " to ", indices[length(indices)])
  str(summarized_df)

  list(summarized_df)  # Return as list to avoid unintended simplification
}

# Flatten the results_list and ensure each element is a data frame
flattened_results <- map_dfr(results_list, ~ as.data.frame(.x))

# Check the structure of the combined result
str(flattened_results)

# Convert to sf object if applicable
if ("geometry" %in% colnames(flattened_results)) {
  spatial_fdis_plant_grid <- st_as_sf(flattened_results)
} else {
  spatial_fdis_plant_grid <- flattened_results
}

# Check the structure of the final output
str(spatial_fdis_plant_grid)

```{r}
plantgridFDisTA_5km <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_plant_grid, aes(fill = fdis_value), color = NA) +
    labs(fill = "FDis") +
    ggtitle("5 km") +
    scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,6.5)) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12), plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 45, hjust = 1))
plantgridFDisTA_5km
```

```{r}
plant_cellFDis_5km <- spatial_fdis_plant_grid
```

```{r}
# add title to maps
(plantgridFDisTA_5km <- plantgridFDisTA_5km +
    ggtitle("5 km")
)
ggsave("plantgridFDisTA_5km.png", plantgridFDisTA_5km, path = figure_path)
```

# Multi panel plots of FD maps
## Make sure all plots for the same taxa have the same legend scale
```{r}
plant_plot_legend <- function(plot){
  plot + scale_fill_distiller(palette = "Greens", direction = 1,limits = c(0, 6.5),
                              oob = scales::squish, breaks = seq(0, 6.5, by = 1)) +
    theme(legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10)) +
      annotation_scale(location = "bl", width_hint = 0.5)
}
frugivore_plot_legend <- function(plot){
  plot + scale_fill_distiller(palette = "YlOrBr", direction = 1,limits = c(0, 8.5),
                              oob = scales::squish, breaks = seq(0, 8.5, by = 1)) +
    theme(legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10)) +
      annotation_scale(location = "bl", width_hint = 0.5)
}
```

```{r}
(plantgridFDisTA_100km <- plant_plot_legend(plantgridFDisTA_100km))
(plantgridFDisTA_75km <- plant_plot_legend(plantgridFDisTA_75km))
(plantgridFDisTA_50km <- plant_plot_legend(plantgridFDisTA_50km))
(plantgridFDisTA_25km <- plant_plot_legend(plantgridFDisTA_25km))
(plantgridFDisTA_10km <- plant_plot_legend(plantgridFDisTA_10km))
(plantgridFDisTA_5km <- plant_plot_legend(plantgridFDisTA_5km))
```
